---
title: "ES193DS-generativeart_project"
author: "Michelle Shteynberg"
date: "2024-05-31"
format: html 
editor: visual
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

link to forked repository:
https://github.com/michelle-shteynberg/ES193DS-generativeart.git

### Part 2a. Workshop (structured only)

#### Introductory exercises

Read and complete exercises from “Get Started” ([link](https://art-from-code.netlify.app/day-1/session-1/)). Annotate your code to demonstrate that you know how it works.

#### 1. Exercise 1
```{r}
library(ggplot2)
library(tibble)

mpg |>
  ggplot(aes(displ, hwy, colour = cyl)) + # setting up the aesthetics, color based on cylinder
  geom_point(show.legend = FALSE, size = 6) + # first layer of points, different shades of blue
  geom_point(show.legend = FALSE, size = 2, colour = "pink2") + # second layer of points, pink
  coord_flip() +  # flipping x and y axes
  theme_void() + # minimal theme
  scale_color_gradient() # adding colors as a gradient
  
```

#### 2. Exercise 2
```{r}

```

#### 3. Exercise 3
```{r}

```

#### 4. Exercise 4

#### Art of your choice: tricks

Choose one: spatial noise tricks, polygon tricks, shading tricks  
Complete all exercises. Again, annotate your code.  

```{r}




```

#### Art of your choice: iteration, tiles, pixels  

Choose one: iterated function systems, tiles and tessellations, pixel filters.  
Complete all exercises. Again, annotate your code.

```{r}




```

#### Creating 3 art pieces

Create 3 art pieces that combines elements of what you learned from parts b-d. For each piece, write an accompanying caption that includes:  

- The title of your piece  
- The date you created it  
- A 2-3 sentence description of your inspiration  
- A 2-3 sentence description of what _specific components_ from each part (b-d) you drew from to create the piece

```{r}




```

```{r}





```

```{r}




```

#### Summary

lastly: a 8-10 sentence summary at the end about your process: What was new for you? What was familiar? What did you learn? How did following the exercises/tutorials go for you?

**enter answer here**